---
title: "predictors_compilation"
output: html_document
date: "2024-07-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Script background

The purpose of this script is to combine 2007 and 2012 nutrient inventories and apply cyanobacteria and microcystin estimator models to the \~125,000 lakes in the data set.

```{r}
# load packages

install.packages("lakemorpho")
install.packages("elevatr")
install.packages("raster")
install_github("USEPA/StreamCatTools", build_vignettes=FALSE, auth_token= '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', force = TRUE)

library(devtools)
library(dplyr)
library(stars)
library(nhdplusTools)
library(tidyverse)
library(tidyr)
library(sf)
library(tigris)
library(StreamCatTools)
library(ggplot2)
library(spmodel)
library(elevatr)
library(lakemorpho)
library(raster)
library(corrplot)
library(remotes)

```

#### Function Development

Establish functions required to load, organize, and clean data

```{r}

# Ensure the files combined as intended, checking for the number of repeated COMIDs
verify <- function(df){
  print(which(df$wbCOMID == "487"))
}

# Visually cleaner way to bind rows, more reproducible
combines <- function(x, y, z){
  x |>
    bind_rows(y) |>
    bind_rows(z)
}

# Aggregating and averaging data across binded data sets
mean_group <- function(df){
  df |>
    group_by(across(wbCOMID)) |>
    summarise(across(where(is.numeric), mean))
}

# Combining variables to main predictor data set, removing repeated columns and merging by COMID
incorp <- function(df){
  df <- subset(df, select = -c(CatAreaSqKm,WsAreaSqKm,CatPctFull,WsPctFull,inStreamCat))
  merge(PredDataMas, df, by = 'COMID')
}

```

### Load Nutrient Data

Nutrient data sourced from Meredith Brehob and Robert Sabo

```{r}

PredData07_05Ws <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/PredData07_05Ws.csv")
PredData07_07Ws <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/PredData07_05Ws.csv")
PredData07_10Ws <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/PredData07_10Ws.csv")
PredData12_05Ws <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/PredData12_05Ws.csv")
PredData12_07Ws <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/PredData12_07Ws.csv")
PredData12_10Ws <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/PredData12_10Ws.csv")

```

### Combine Year Datasets

2007 and 2012 data 

```{r}

PredData2007 <- combines(PredData07_05Ws, PredData07_07Ws, PredData07_10Ws)
verify(PredData2007)

PredData2007 <- mean_group(PredData2007)
verify(PredData2007)

# 2012

PredData2012 <- combines(PredData12_05Ws, PredData12_07Ws, PredData12_10Ws)
verify(PredData2012)

PredData2012 <- mean_group(PredData2012)
verify(PredData2012)

# Create the Master Dataset ------------------------------------------------

PredDataMas <- PredData2007 |>
  bind_rows(PredData2012)
verify(PredDataMas)

PredDataMas <- mean_group(PredDataMas)
verify(PredDataMas)

# rename the COMID variable to match other data sets
names(PredDataMas)[names(PredDataMas) == "wbCOMID"] <- "COMID"

# rename the Runoff variable to clarify it was derived from the nutrient inventory data
names(PredDataMas)[names(PredDataMas) == "RunoffWs"] <- "Runoff.Nutr"

head(PredDataMas)

```

### Combine Data Sets

Supporting variables sourced from StreamCat Data developed by Marc Weber and Ryan Hill

```{r}

# pesticide data from 1997 StreamCat
pesticides <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/lakecat-metrics-melanie/Pesticides97.csv")
PredDataMas <- incorp(pesticides)

# base flow index derived from Streamcat
BFI <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/lakecat-metrics-melanie/BFI.csv")
names(BFI)[names(BFI) == "BFIWs"] <- "BFIWs.Str" # Renmaed for clarity
PredDataMas <- incorp(BFI)

# PRISM data derived from Stream Cat
PRISM <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/lakecat-metrics-melanie/PRISM_1991_2020.csv")
PredDataMas <- incorp(PRISM)

# Runoff data derived from Stream Cat
runoff <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/lakecat-metrics-melanie/runoff.csv")

names(runoff)[names(runoff) == "RunoffWs"] <- "Runoff.Str"
runoff <- subset(runoff, select = -c(RunoffCat))
PredDataMas <- incorp(runoff)

# Lithological Nitrogen, Stream Cat
rockN <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/PredData/lakecat-metrics-melanie/RockN.csv")
PredDataMas <- incorp(rockN)

# Ecoregions data, derived from NLA
eco3 <- read.csv("C:/Users/mreyno04/OneDrive - Environmental Protection Agency (EPA)/Profile/REPOS/LakePredData/NLA_SampleFrame_L3L4Ecoregions 1.csv")
names(eco3)[names(eco3) == "comid"] <- "COMID"

PredDataMas <- merge(PredDataMas, eco3, by = 'COMID')

# Renaming variables for the merge
names(PredDataMas)[names(PredDataMas) == "BFIWs"] <- "BFIWs.Nutr" # BFIWs 2007-2012 data compiled

# Remove excess column
PredDataMas = subset(PredDataMas, select = -c(X))

```

### Land Cover Data

```{r}

# function set up to pull desired land cover data based on COMIDs
get_nlcd <- function(coms){
  lc_get_data(metric = 'PctWdWet2016, PctUrbMd2016, PctUrbLo2016, PctUrbHi2016,
                          PctMxFst2016, PctCrop2016, PctHay2016, PctDecid2016,
                          PctConif2016, PctUrbOp2016, PctHbWet2016',
                    aoi='watershed',
                    comid = coms,
                    showAreaSqKm = TRUE)
}

# dividing COMIDs from Predictor Dataset into chunks to prevent data overload
chunks <- split(PredDataMas$COMID, ceiling(seq_along(PredDataMas$COMID) / 10000))

# apllying function to exsiting COMIDs
ncldMas2 <- do.call(rbind, lapply(chunks, get_nlcd))

# merge land cover data and main predictor set
PredDataMas <- merge(PredDataMas, ncldMas2, by = 'COMID')

# remove and rename overlapping columns
PredDataMas <- subset(PredDataMas, select = -c(WSAREASQKM,WsAreaSqKm.y))
names(PredDataMas)[names(PredDataMas) == "WsAreaSqKm.x"] <- "WsAreaSqKm"

# variables for the model

# agr_ws
# NLCD agricultural land cover including hay and cultivated crop. Data are paired as follows: NLA 2007 - NLCD 2006, NLA 2012 - NLCD 2011, NLA 2017 - NLCD 2016 (percent)

PredDataMas$agr_ws <- PredDataMas$PCTCROP2016WS + PredDataMas$PCTHAY2016WS

# dev_ws
# NLCD developed land cover including open spaces, low, medium, and high intensity. Data are paired as follows: NLA 2007 - NLCD 2006, NLA 2012 - NLCD 2011, NLA 2017 - NLCD 2016 (percent)

PredDataMas$dev_ws <- PredDataMas$PCTURBLO2016WS + PredDataMas$PCTURBMD2016WS + PredDataMas$PCTURBOP2016WS + PredDataMas$PCTURBHI2016WS

# fst_ws
# NLCD forested land cover including coniferous, deciduous, and mixed forest. Data are paired as follows: NLA 2007 - NLCD 2006, NLA 2012 - NLCD 2011, NLA 2017 - NLCD 2016 (percent)

PredDataMas$fst_ws <- PredDataMas$PCTMXFST2016WS + PredDataMas$PCTDECID2016WS + PredDataMas$PCTCONIF2016WS

# wet_ws
# NLCD wetland land cover including herbaceous and woody wetlands. Data are paired as follows: NLA 2007 - NLCD 2006, NLA 2012 - NLCD 2011, NLA 2017 - NLCD 2016 (percent)

PredDataMas$wet_ws <- PredDataMas$PCTWDWET2016WS + PredDataMas$PCTHBWET2016WS

```

### Waterbody Data

```{r}

# load in waterbody data 
loc <- "O:/PRIV/CPHEA/PESD/COR/CORFILES/Geospatial_Library_Resource/Physical/HYDROLOGY/NHDPlusV21/NHDPlusNationalData/NHDPlusV21_National_Seamless_Flattened_Lower48.gdb"

wbd <- sf::st_read(dsn = loc, layer = "NHDWaterbody") |>
  st_transform(5072)

# create wbd copy file with only desired COMIDs for easier computation
wbd_copy <- subset(wbd, COMID %in% PredDataMas$COMID)

```

### Data Cleaning

```{r}

# remove unnecessary eco-regions data
PredDataMas <- subset(PredDataMas, select = -c(US_L4CODE, US_L4NAME, US_L3CODE, US_L3NAME, NA_L3CODE, NA_L3NAME,NA_L2CODE, NA_L2NAME, NA_L1CODE, NA_L1NAME, nla07_sf, nla12_sf, nla17_sf))

```

### Nutrient Data

```{r}

# load files from folder
data_dir <- "O:/PRIV/CPHEA/PESD/COR/CORFILES/Geospatial_Library_Projects/AmaliaHandler/Allocation_and_Accumulation/Final_LakeCat_OnNetwork/"
csv_files <- fs::dir_ls(data_dir)

# compile files into a single data frame
nutrMas <- csv_files |>
  map_dfr(read_csv)

# ensure there's no repetition
print(which(nutrMas$COMID == "487"))

# remove unnecessary columns
nutrMas <- nutrMas |> dplyr::select(-contains(c('Cat')))

# aggregate nutrients across years

# N livestock waste
nutrMas$N_livestock.Waste_kg_Ag <- rowMeans(nutrMas[,c('N_Livestock.Waste_kg_Ag_2002Ws', 'N_Livestock.Waste_kg_Ag_2007Ws', 'N_Livestock.Waste_kg_Ag_2012Ws')], na.rm = TRUE)
nutrMas <- subset(nutrMas, select = -c(N_Livestock.Waste_kg_Ag_2002Ws, N_Livestock.Waste_kg_Ag_2007Ws, N_Livestock.Waste_kg_Ag_2012Ws))

# P livestock waste
nutrMas$P_livestock_Waste_kg_Ag <- rowMeans(nutrMas[,c('P_livestock_Waste_kg_Ag_2002Ws', 'P_livestock_Waste_kg_Ag_2007Ws', 'P_livestock_Waste_kg_Ag_2012Ws')], na.rm = TRUE)
nutrMas <- subset(nutrMas, select = -c(P_livestock_Waste_kg_Ag_2002Ws, P_livestock_Waste_kg_Ag_2007Ws, P_livestock_Waste_kg_Ag_2012Ws))

# P ag fertilizer
nutrMas$P_f_fertilizer_kg_Ag <- rowMeans(nutrMas[,c('P_f_fertilizer_kg_Ag_2002Ws', 'P_f_fertilizer_kg_Ag_2007Ws', 'P_f_fertilizer_kg_Ag_2012Ws')], na.rm = TRUE)
nutrMas <- subset(nutrMas, select = -c(P_f_fertilizer_kg_Ag_2002Ws, P_f_fertilizer_kg_Ag_2007Ws, P_f_fertilizer_kg_Ag_2012Ws))

# N human waste
nutrMas$N_Human_Waste_kg_Urb <- rowMeans(nutrMas[,c('N_Human_Waste_kg_Urb_2002Ws', 'N_Human_Waste_kg_Urb_2007Ws', 'N_Human_Waste_kg_Urb_2012Ws')], na.rm = TRUE)
nutrMas <- subset(nutrMas, select = -c(N_Human_Waste_kg_Urb_2002Ws, N_Human_Waste_kg_Urb_2007Ws, N_Human_Waste_kg_Urb_2012Ws))

# P human waste
nutrMas$P_Human_Waste_kg_Urb <- rowMeans(nutrMas[,c('P_human_waste_kg_Urb_2002Ws', 'P_human_waste_kg_Urb_2007Ws', 'P_human_waste_kg_Urb_2012Ws')], na.rm = TRUE)
nutrMas <- subset(nutrMas, select = -c(P_human_waste_kg_Urb_2002Ws, P_human_waste_kg_Urb_2007Ws, P_human_waste_kg_Urb_2012Ws))

# N urban fertilizer
nutrMas$N_Fert_Urban_kg_Urb <- rowMeans(nutrMas[,c('N_Fert_Urban_kg_Urb_2002Ws', 'N_Fert_Urban_kg_Urb_2007Ws', 'N_Fert_Urban_kg_Urb_2012Ws')], na.rm = TRUE)
nutrMas <- subset(nutrMas, select = -c(N_Fert_Urban_kg_Urb_2002Ws, N_Fert_Urban_kg_Urb_2007Ws, N_Fert_Urban_kg_Urb_2012Ws))

# P not from fertilizer
nutrMas$P_nf_fertilizer_kg_Urb <- rowMeans(nutrMas[,c('P_nf_fertilizer_kg_Urb_2002Ws', 'P_nf_fertilizer_kg_Urb_2007Ws', 'P_nf_fertilizer_kg_Urb_2012Ws')], na.rm = TRUE)
nutrMas <- subset(nutrMas, select = -c(P_nf_fertilizer_kg_Urb_2002Ws, P_nf_fertilizer_kg_Urb_2007Ws, P_nf_fertilizer_kg_Urb_2012Ws))

# rename CBNF variable
names(nutrMas)[names(nutrMas) == 'N_CBNF_kg_Ag_2002Ws'] <- 'N_CBNF_kg_Ag'

# check all column names
colnames(nutrMas)

# combine nutrient data with main predictor data set
PredDataMas <- merge(PredDataMas, nutrMas, by = 'COMID')

```
